---
title: "GLM MST"
author: "Dimitrios Zacharatos"
date: "`r Sys.Date()`"
output:
  html_document:
    css: styles.css
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: readable
vignette: >
  %\VignetteIndexEntry{GLM MST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options,echo=FALSE,warning=FALSE,message=FALSE}
options(width=1000)
hook_output <- function(x, options) {
  paste0('<pre class="r-output">', knitr::knit_print(x), '</pre>')
}
knitr::knit_hooks$set(output = hook_output)
knitr::opts_chunk$set(echo=TRUE)
options(future.show.progress = FALSE)
```

```{r library,message=FALSE,warning=FALSE}
library(rwf)
library(mstR)
```

## Description

This vignette demonstrates Multi-Stage Testing (MST) using the `mstR` package alongside helper functions from `rwf`. MST is an adaptive testing methodology that sits between traditional linear tests and fully adaptive Computerised Adaptive Testing (CAT). Instead of selecting items one at a time, MST groups items into **modules** and selects the next module based on the examinee's performance on the previous one.

A typical MST design has multiple **stages** (e.g., stage 1 → stage 2 → stage 3) and multiple **panels** (parallel test forms). Within each stage, an examinee is routed to an easy, medium, or hard module depending on their estimated ability (θ). This balances measurement precision with practical constraints such as content balancing and test security.

Key concepts used throughout this vignette:

- **Item bank**: a pool of items described by IRT parameters (here, the 2-Parameter Logistic model — discrimination `a` and difficulty `b`).
- **Modules**: subsets of items from the bank assigned to specific positions in the MST design.
- **Transition matrix**: defines which modules can follow which — routing rules between stages.
- **Cutoff matrix**: θ or raw-score thresholds that determine which module an examinee is routed to next.
- **θ (theta)**: the latent ability estimate on the standard normal scale.

Installation instructions for rwf can be found on the [rwf GitHub repository](https://github.com/sedzinfo/rwf)

## Basic Examples

This section constructs a small 1-3-2 MST design by hand and demonstrates the two core routing functions: `startModule` and `nextModule`.

**Item bank construction**: Six groups of items are generated using `genDichoMatrix` under the 2PL model. The difficulty priors (`bPrior`) are shifted so that items in different modules cover different regions of the ability scale — negative difficulties for easy modules and positive difficulties for hard modules.

**Module matrix** (`modules`): A 47×6 binary matrix assigning each of the 47 items to one of six modules. Each column is a module; a `1` indicates that item belongs to that module. `colSums(modules)` confirms the expected module sizes (8, 6, 6, 9, 9, 9).

**Transition matrix** (`trans`): A 6×6 binary matrix defining allowed routing paths. Here, module 1 (stage 1) can route to modules 2 or 3 (stage 2); module 2 routes to modules 4 or 5; module 3 routes to modules 5 or 6. Module 5 is reachable from both stage-2 modules, acting as the medium-difficulty path.

**`startModule`**: Returns which module to administer first given the item bank, module structure, and transition matrix. The optional `fixModule` argument (commented out) would force a specific starting module regardless of the design.

**`thetaEst`**: Estimates θ from the response pattern `x` on the first 8 items using Maximum Likelihood (ML). This estimate is then used to determine the next module.

**`nextModule` with θ cutoffs** (`cut`): Routes to the next module based on whether the ML θ estimate falls below or above the defined thresholds (e.g., θ < 0 → module 2; θ ≥ 0 → module 3). Two calls show routing for the estimated θ and for a fixed θ of −0.2.

**`nextModule` with score cutoffs** (`cut.score`): Routes based on raw number-correct scores rather than θ estimates. This is useful when IRT scoring is not available or when content-balancing rules are score-based.

```{r excamples0,fig.width=8,fig.height=8,echo=TRUE,message=FALSE,warning=FALSE,results='hide',dev='png',fig.retina=2}
it.MST <- rbind(genDichoMatrix(8, model = "2PL"), 
                genDichoMatrix(6, model = "2PL",bPrior = c("norm", -1, 1)),
                genDichoMatrix(6, model = "2PL",bPrior = c("norm", 1, 1)),
                genDichoMatrix(9, model = "2PL",bPrior = c("norm", -2, 1)),
                genDichoMatrix(9, model = "2PL",bPrior = c("norm", 0, 1)),
                genDichoMatrix(9, model = "2PL",bPrior = c("norm", 2, 1)))
it.MST <- as.matrix(it.MST)

modules <- matrix(0, 47, 6)
modules[1:8, 1] <- 1
modules[9:14, 2] <- 1
modules[15:20, 3] <- 1
modules[21:29, 4] <- 1
modules[30:38, 5] <- 1
modules[39:47, 6] <- 1

colSums(modules)

trans <- matrix(0, 6, 6)
trans[1, 2:3] <- 1
trans[2, 4:5] <- 1
trans[3, 5:6] <- 1

startModule(itemBank = it.MST, modules = modules, transMatrix = trans)
# startModule(itemBank = it.MST, modules = modules, transMatrix = trans, fixModule = 3)

x <- c(1, 1, 1, 1, 0, 0, 0, 0)
th<-mstR::thetaEst(it.MST[1:8, ], x, method = "ML")

cut <- matrix(NA, 6, 2)
cut[2,] <- c(-Inf, 0)
cut[3,] <- c(0, Inf)
cut[4,] <- c(-Inf, -1)
cut[5,] <- c(-1, 1)
cut[6,] <- c(1, Inf)

nextModule(itemBank = it.MST, modules = modules,transMatrix = trans, current.module = 1, out = 1:8, cutoff = cut, theta = th)
nextModule(itemBank = it.MST, modules = modules, transMatrix = trans, current.module = 1, out = 1:8, cutoff = cut, theta = -0.2)

cut.score <- matrix(NA, 6, 2)
cut.score [2,] <- c(0, 5)
cut.score [3,] <- c(5, 8)
cut.score [4,] <- c(0, 5)
cut.score [5,] <- c(5, 10)
cut.score [6,] <- c(10, 14)

nextModule(itemBank = it.MST, modules = modules, transMatrix = trans, current.module = 3, out = c(1:8, 15:20), cutoff = cut.score,theta = th)

```


## Simulation Examples

This section runs full MST simulations using `randomMST` and compares MST performance against CAT using `randomCAT`. A 100-item bank with five 2PL items is used, assembled into six modules of 15 items each across a 1-2-2-1 branching design.

**Item bank** (`it.2PL`): Five 2PL items with discrimination values from 0.8 to 1.2 and difficulties spanning −0.8 to 1.1, recycled to fill a 100-item bank. Parameters `c = 0` and `d = 1` indicate no guessing and no slipping.

**Module membership** (A1–F1): Six pre-defined sets of 15 item indices each, assembled into `module1`. A1 is the stage-1 routing module; B1/C1 are stage-2 easy/hard modules; D1/E1/F1 are the three stage-3 modules.

**Simulation 1 — MFI module selection** (`mst.ex1`): A single MST run at true θ = 0 using ML scoring and Maximum Fisher Information (MFI) for module selection. The two `Ii()` calls confirm the routing decision by comparing total Fisher information in module 2 vs. module 3 at the stage-1 θ estimate — the module with higher information should be selected. `plot(mst.ex1)` draws the routing path taken through the MST design.

**Simulation 2 — θ cutoff routing** (`mst.ex2`): Routes modules using θ thresholds (θ < 0 → easy; 0 ≤ θ ≤ 0.3 → medium; θ > 0.3 → hard). Run at true θ = 1 to verify that the design routes to the harder modules.

**Simulation 3 — score cutoff routing** (`mst.ex3`): Routes using raw number-correct score thresholds rather than θ estimates. The scoring method is set to `"score"` for both routing and final estimation — useful when IRT calibration is not available at administration time.

**ASB and RMSE across the ability range**: A large-scale simulation (1000 replications × 9 ability levels from −2 to +2 × 6 module configurations) estimates two accuracy metrics across the full latent trait range:

- **ASB (Average Signed Bias)**: mean difference between estimated and true θ — positive values indicate systematic overestimation, negative indicate underestimation.
- **RMSE (Root Mean Square Error)**: overall estimation accuracy combining both bias and variance. Lower values indicate a more precise test design.

**MST vs. CAT comparison**: The final block compares MST (BM scoring, MFI module selection) against CAT (45-item stopping rule, BM scoring, MFI item selection, randomesque = 5 for item exposure control) across 9000 simulated examinees spanning θ from −2 to +2. Both `thFinal` (final ability estimate) and `seFinal` (standard error of measurement) are stored in `RES` for comparison of precision between the two adaptive testing formats.

```{r excamples1,fig.width=8,fig.height=8,echo=TRUE,message=FALSE,warning=FALSE,results='hide',dev='png',fig.retina=2}
A1 <- c(7, 13, 14, 15, 22, 28, 34, 38, 51, 59, 70, 84, 85, 87, 89)
B1 <- c(4, 12, 19, 27, 29, 39, 43, 60, 69, 71, 72, 73, 75, 90, 97)
C1 <- c(32, 35, 36, 45, 52, 57, 61, 63, 66, 78, 81,82, 88, 92, 100)
D1 <- c(1, 2, 3, 6, 18, 21, 41, 48, 49, 50, 54, 56, 58, 65, 77)
E1 <- c(5, 16, 24, 25, 26, 37, 40, 44, 53, 68, 74, 80, 86, 91, 93)
F1 <- c(9, 11, 31, 46, 47, 55, 64, 67, 76, 79, 83, 94, 95, 98, 99)

a <- c(0.8, 0.9, 1.0, 1.1, 1.2)
b <- c(0.0, -0.5, 1.1, -0.8, 0.7)
it.2PL <- cbind(a, b, c = 0, d = 1)

module1 <- matrix(0, 100, 6)
module1[A1, 1] <- module1[B1, 2] <- module1[C1, 3] <- module1[D1, 4] <- module1[E1, 5] <- module1[F1, 6] <- 1

trans <- matrix(0, 6, 6)
trans[1, 2:3] <- trans[2, 4:5] <- trans[3, 5:6] <- 1

start <- list(theta = 0)

test <- list(method = "ML", moduleSelect = "MFI")
final <- list(method = "ML")
mst.ex1 <- randomMST(trueTheta = 0, itemBank = it.2PL, modules = module1, transMatrix = trans, genSeed = 1, start = start, test = test, final = final)

sum(Ii(0.204, it.2PL[module1[, 2] == 1, ])$Ii)
sum(Ii(0.204, it.2PL[module1[, 3] == 1, ])$Ii)

plot(mst.ex1, border.col = "grey", arrow.col = "grey")

cut <- matrix(NA, 6, 2)
cut[2,] <- c(-Inf, 0)
cut[3,] <- c(0, Inf)
cut[4,] <- c(-Inf,-0.3)
cut[5,] <- c(-0.3, 0.3)
cut[6,] <- c(0.3, Inf)

test2 <- list(method = "ML", cutoff = cut)
mst.ex2 <- randomMST(trueTheta = 1, itemBank = it.2PL, modules = module1, transMatrix = trans, genSeed = 1, start = start, test = test2, final = final)

cut2 <- matrix(NA, 6, 2)
cut2[2,] <- c(-Inf, 7.5)
cut2[3,] <- c(7.5, Inf)
cut2[4,] <- c(-Inf,9.5)
cut2[5,] <- c(9.5, 20.5)
cut2[6,] <- c(20.5, Inf)

test3 <- list(method = "score", cutoff = cut2)
final3 <- list(method = "score")
mst.ex3 <- randomMST(trueTheta = 1, itemBank = it.2PL, modules = module1, transMatrix = trans, genSeed = 1, start = start, test = test3, final = final)

s <- seq(-2, 2, 0.5)
TH <- rep(s, each = 1000)
RES <- matrix(NA, length(TH), 6)
start <- list(theta = 0)
test <- list(method = "BM", moduleSelect = "MFI")
final <- list(method = "ML")

for (j in 1:6) { 
  mod <- get(paste("module", j, sep = ""))
  for (i in 1:length(TH)){
    prov <- randomMST(trueTheta = TH[i],
                      itemBank = it.2PL, modules = mod,
                      transMatrix = trans, genSeed = i, start = start,
                      test = test, final = final)
    RES[i, j] <- prov$thFinal
  }}

res.ASB <- res.RMSE <- matrix(NA, length(s), 6)
ASB <- function(t, th) mean(t-th)
RMSE <- function(t, th) sqrt(mean((t-th)^2))

for (i in 1:length(s)){
  for (j in 1:6){
    ind <- which(TH == s[i])
    res.ASB[i, j] <- ASB(RES[ind, j], s[i])
    res.RMSE[i, j] <- RMSE(RES[ind, j], s[i])
  }}

s <- seq(-2, 2, 0.5)
TH <- rep(s, each = 1000)
DATA <- genPattern(TH, it.2PL, seed = 1)
RES <- matrix(NA, length(TH), 4)

start.mst <- list(theta = 0)
start.cat <- list(theta = 0, randomesque = 5)
test.cat <- list(method = "BM", itemSelect = "MFI", randomesque = 5)
test.mst <- list(method = "BM", moduleSelect = "MFI")
stop <- list(rule = "length", thr = 45)
final <- list(method = "BM")

for (i in 1:length(TH)){
  pr <- randomCAT(trueTheta = TH[i],
                  itemBank = it.2PL, genSeed = i, start = start.cat,
                  test = test.cat, stop = stop, final = final)
  RES[i,1] <- pr$thFinal
  RES[i,2] <- pr$seFinal
  pr <- randomMST(trueTheta = TH[i],
                  itemBank = it.2PL, genSeed = i, modules = module1,
                  transMatrix = trans, start = start.mst,
                  test = test.mst, final = final)
  RES[i,3] <- pr$thFinal
  RES[i,4] <- pr$seFinal
}


```


























