---
title: "FUNCTIONS TRAIN TEST"
author: "Dimitrios Zacharatos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FUNCTIONS_TRAIN_TEST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
/* Styling for input code */
pre { 
  background-color: #f4f4f4; /* Light gray background for input */
  border-left: 5px solid #007ACC; /* Blue border */
  padding: 10px;
}

/* Styling for output */
pre.r-output {
  background-color: #e8f5e9; /* Light green background for output */
  border-left: 5px solid #388E3C; /* Green border */
  padding: 10px;
}
</style>

This shows the output of useful functions for training and testing from workingfunctions.  
Installation instructions for workingfunctions can be found [here](https://github.com/sedzinfo/workingfunctions)

```{r setup, include=FALSE}
library(workingfunctions)
knitr::opts_chunk$set(echo=TRUE)
options(width=1000)
options(future.show.progress=FALSE)
# Custom hook for output styling
hook_output <- function(x, options) {
  paste0('<pre class="r-output">', knitr::knit_print(x), '</pre>')
}
knitr::knit_hooks$set(output = hook_output)

```

## Plot ROC Curve
```{r plot_roc,fig.width=8,fig.height=8,echo=TRUE,message=FALSE,warning=FALSE,results='hide',dev='png',fig.retina=2}
observed<-round(abs(rnorm(100,m=0,sd=.5)))
predicted<-abs(rnorm(100,m=0,sd=.5))
plot_roc(observed=observed,predicted=predicted)
df1<-data.frame(matrix(.999,ncol=2,nrow=2))
correlation_martix<-as.matrix(df1)
diag(correlation_martix)<-1
df1<-generate_correlation_matrix(correlation_martix,nrows=1000)
df1$X1<-ifelse(abs(df1$X1)<1,0,1)
df1$X2<-abs(df1$X2)
df1$X2<-(df1$X2-min(df1$X2))/(max(df1$X2)-min(df1$X2))
plot_roc(observed=round(abs(df1$X1),0),predicted=abs(df1$X2))
```

## Plot Confusion Matrix
```{r plot_confusion,fig.width=8,fig.height=8,dev='png',fig.retina=2}
plot_confusion(observed=c(1,2,3,1,2,3),predicted=c(1,2,3,1,2,3))
observed<-c(rep("male",10),rep("female",10),"male","male")
predicted<-c(rep("male",10),rep("female",10),"female","female")
plot_confusion(observed=observed,predicted=predicted)
```

## Plot Separability
```{r plot_separability,fig.width=8,fig.height=8,dev='png',fig.retina=2}
df1<-data.frame(matrix(.999,ncol=2,nrow=2))
correlation_martix<-as.matrix(df1)
diag(correlation_martix)<-1
df1<-generate_correlation_matrix(correlation_martix,nrows=1000)
df1$X1<-ifelse(abs(df1$X1)<1,0,1)
df1$X2<-abs(df1$X2)
df1$X2<-(df1$X2-min(df1$X2))/(max(df1$X2)-min(df1$X2))
plot_separability(observed=round(abs(df1$X1),0),predicted=abs(df1$X2))
```

## Plot Confusion Matrix Performance
```{r plot_confusion_performance,fig.width=9,fig.height=8,dev='png',fig.retina=2}
df<-data.frame(matrix(.999,ncol=2,nrow=2))
correlation_martix<-as.matrix(df)
diag(correlation_martix)<-1
df<-generate_correlation_matrix(correlation_martix,nrows=1000)
df$X1<-ifelse(abs(df$X1)<1,0,1)
df$X2<-abs(df$X2)
df$X2<-(df$X2-min(df$X2))/(max(df$X2)-min(df$X2))
result_confusion_performance(observed=round(abs(df$X1),0),
                             predicted=abs(df$X2),
                             step=.01)
result_confusion_performance(observed=c(1,2,3,1,2,3),predicted=abs(rnorm(6,0,sd=.1)))
```

## KFOLD
```{r kfold}
infert_formula<-as.formula(factor(case)~age+parity+education+spontaneous+induced)
result<-k_fold(infert,k=10,model_formula=infert_formula)
model_formula<-as.formula(mpg~cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb)
result<-k_fold(mtcars,k=2,model_formula=model_formula)
result
```

## KSAMPLE
```{r ksample}
infert_formula<-as.formula(factor(case)~age+parity+education+spontaneous+induced)
result<-k_sample(df=infert,k=10,model_formula=infert_formula)
model_formula<-as.formula(mpg~cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb)
result<-k_sample(df=mtcars,k=10,model_formula=model_formula)
result
```

## Scaling
```{r scaling}
recode_scale_dummy(infert)
```

## Confusion Matrix for Classification
```{r confusion_classification}
confusion(observed=c(1,2,3,4,5,10),predicted=c(1,2,3,4,5,11))
confusion(observed=c(1,2,2,2,2),predicted=c(1,1,2,2,2))
```

## Proportion Accutate
```{r proportion_accurate}
proportion_accurate(observed=c(1,2,3,4,5,10),predicted=c(1,2,3,4,5,11))
```

## Confusion Matrix Percent
```{r confusion_percent}
observed<-factor(round(rnorm(10000,m=10,sd=1)))
predicted<-factor(round(rnorm(10000,m=10,sd=1)))
confusion_matrix_percent(observed,predicted)
```







