---
title: "GLM_EFA"
author: "Dimitrios Zacharatos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GLM_EFA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  <style> 
  body { max-width: 95%; margin: 0 auto; padding: 0; } 
  .container { width: 100%; max-width: none; } 
  /* Example for wider tables */ 
  table { width: 80%; table-layout: fixed; } 
  th, td { padding: 8px; text-align: left; } 
  </style>
  
  This shows the output of several EFA functions from workingfunctions.  
  Installation instructions of workingfunctions can be found here https://github.com/sedzinfo/workingfunctions
  
```{r setup}
library(workingfunctions)
options(width=1000)
```

# Plot Loadings
```{r, plot_loadings,fig.width=10,fig.height=5}
model<-psych::fa(mtcars,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
plot_loadings(model=model,matrix_type="structure")
plot_loadings(model=model,matrix_type="pattern")
cm<-matrix(c(1,.8,.8,.1,.1,.1,
             .8,1,.8,.1,.1,.1,
             .8,.8,1,.1,.1,.1,
             .1,.1,.1,1,.8,.8,
             .1,.1,.1,.8,1,.8,
             .1,.1,.1,.8,.8,1),
           ncol=6,nrow=6)
df1<-generate_correlation_matrix(cm,nrows=10000)
model1<-psych::fa(df1,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
plot_loadings(model=model1,matrix_type="pattern",base_size=10)
cm<-matrix(c(1,.1,.1,.1,.1,.1,
             .1,1,.1,.1,.1,.1,
             .1,.1,1,.1,.1,.1,
             .1,.1,.1,1,.8,.8,
             .1,.1,.1,.8,1,.8,
             .1,.1,.1,.8,.8,1),
           ncol=6,nrow=6)
df1<-generate_correlation_matrix(cm,nrows=10000)
model2<-psych::fa(df1,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
plot_loadings(model=model2,matrix_type="pattern",base_size=10)
cm<-matrix(c(1,.01,.01,.01,.01,.01,
             .01,1,.01,.01,.01,.01,
             .01,.01,1,.01,.01,.01,
             .01,.01,.01,1,.01,.01,
             .01,.01,.01,.01,1,.01,
             .01,.01,.01,.01,.01,1),
           ncol=6,nrow=6)
df1<-generate_correlation_matrix(cm,nrows=10000)
model3<-psych::fa(df1,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
plot_loadings(model=model3,matrix_type="pattern",base_size=10)
```


# Plot Scree
```{r, plot_scree,fig.width=10,fig.height=5}
plot_scree(df=mtcars,title="",base_size=15)
```

# Model Loadings
```{r, model_loadings,fig.width=20,fig.height=10}
model<-psych::fa(mtcars,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
model_loadings(model=model,cut=NULL,matrix_type="pattern")
model_loadings(model=model,cut=0.4,matrix_type="structure")
model_loadings(model=model,cut=0.4,matrix_type="all",sort=FALSE)
```

# Compute Residual
```{r, compute_residual,fig.width=10,fig.height=5}
model<-psych::fa(mtcars,nfactors=2,rotate="oblimin",fm="pa",oblique.scores=TRUE)
compute_residual_stats(model)
```


# Report EFA
```{r, report_efa,fig.width=10,fig.height=5}
model<-psych::fa(mtcars,nfactors=2,rotate="oblimin",fm="minres",oblique.scores=TRUE)
report_efa(model=model,df=mtcars)
```






