---
title: "GLM LINEAR REGRESSION"
author: "Dimitrios Zacharatos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GLM_LINEAR_REGRESSION}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
<style>
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
/* Styling for input code */
pre { 
  background-color: #f4f4f4; /* Light gray background for input */
  border-left: 5px solid #007ACC; /* Blue border */
  padding: 10px;
  font-size: 20px !important;
}
/* Styling for output */
pre.r-output {
  background-color: #e8f5e9; /* Light green background for output */
  border-left: 5px solid #388E3C; /* Green border */
  padding: 10px;
  font-size: 15px !important;
}
</style>

```{r options,echo=FALSE,warning=FALSE,message=FALSE}
options(future.show.progress=FALSE)
options(width=1000)
knitr::opts_chunk$set(echo=TRUE)
hook_output<-function(x, options) {
  paste0('<pre class="r-output">',knitr::knit_print(x),'</pre>')
}
knitr::knit_hooks$set(output=hook_output)
```

```{r load,include=FALSE}
library(workingfunctions)
```

## Description
This shows the output of linear regression functions from workingfunctions.  
Installation instructions of working functions can be found here https://github.com/sedzinfo/workingfunctions

# Plot Scatterplot
```{r, plot_scatterplot,fig.width=8,fig.height=8,message=FALSE,warning=FALSE,results='hide'}
result<-plot_scatterplot(df=mtcars,title="",coord_equal=TRUE,base_size=10)
plot_scatterplot(df=mtcars[,1:2],base_size=10,coord_equal=TRUE,all_orders=FALSE)
plot_scatterplot(df=mtcars[,1:2],base_size=10,coord_equal=FALSE,all_orders=FALSE)
plot_scatterplot(df=mtcars,base_size=10,coord_equal=TRUE,all_orders=FALSE,
                 combinations=data.frame(x=c("mpg","mpg","mpg"),
                                         y=c("cyl","hp","mpg")))
plot_scatterplot(df=mtcars,base_size=10,coord_equal=TRUE,all_orders=TRUE,
                 combinations=data.frame(x=c("mpg"),y=c("cyl")))
x<-rnorm(1000)
y<-x+rnorm(x,sd=.1)
plot_scatterplot(df=data.frame(x,y),title="Random Simulation",coord_equal=TRUE)
df<-data.frame(matrix(-.999,ncol=2,nrow=2))
correlation_martix<-as.matrix(df)
diag(correlation_martix)<-1
df<-generate_correlation_matrix(correlation_martix,nrows=1000)
plot_scatterplot(df,title="Simulation of -.999 Correlation",coord_equal=TRUE,base_size=10)
```

```{r, plot_multiplot,fig.width=20,fig.height=20}
plot_multiplot(plotlist=result[1:12],cols=4)
```

```{r, report_regression,fig.width=8,fig.height=8}
form<-formula(mpg~qsec)
regressionmodel<-lm(form,data=mtcars)
multipleregressionmodel<-lm(mpg~qsec*hp*wt*drat,data=mtcars)
res<-report_regression(model=regressionmodel,plot_diagnostics=TRUE)
res<-report_regression(model=multipleregressionmodel)
res
```
