---
title: "GLM ANOVA"
author: "Dimitrios Zacharatos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GLM ANOVA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
/* Styling for input code */
pre { 
  background-color: #f4f4f4; /* Light gray background for input */
  border-left: 5px solid #007ACC; /* Blue border */
  padding: 10px;
  font-size: 20px !important;
}
/* Styling for output */
pre.r-output {
  background-color: #e8f5e9; /* Light green background for output */
  border-left: 5px solid #388E3C; /* Green border */
  padding: 10px;
  font-size: 15px !important;
}
</style>

```{r options,echo=FALSE,warning=FALSE,message=FALSE}
options(future.show.progress=FALSE)
options(width=1000)
knitr::opts_chunk$set(echo=TRUE)
hook_output<-function(x, options) {
  paste0('<pre class="r-output">',knitr::knit_print(x),'</pre>')
}
knitr::knit_hooks$set(output=hook_output)
```

```{r load,include=FALSE}
library(workingfunctions)
```

## Description
This shows the output of several ANOVA related functions from workingfunctions.  
Installation instructions for workingfunctions can be found [here](https://github.com/sedzinfo/workingfunctions)

## KRUSKALL WALLIS TEST
```{r Kruskal-Wallis,echo=TRUE}
form<-formula(qsec~cyl)
kruskal.test(formula=form,data=mtcars)
rcompanion::epsilonSquared(x=mtcars$qsec,g=mtcars$cyl,group="row",ci=TRUE,conf=0.95,type="perc",R=1000,digits=3)
rstatix::kruskal_effsize(mtcars,form,ci=TRUE,conf.level=0.95,ci.type="perc",nboot=100)
compute_kruskal_wallis_test(formula=form,df=mtcars)
```

## ONE WAY test with SS MS effect size and power
```{r one-way,echo=TRUE}
form<-formula(qsec~cyl)
compute_one_way_test(formula=form,df=mtcars,var.equal=TRUE)
compute_one_way_test(formula=form,df=mtcars,var.equal=FALSE)
oneway.test(formula=form,data=mtcars,var.equal=TRUE)
oneway.test(formula=form,data=mtcars,var.equal=FALSE)
car::Anova(aov(form,data=mtcars),type=2)
model<-lm(form,data=mtcars)
lsr::etaSquared(aov(form,data=mtcars),type=3,anova=TRUE)
sjstats::anova_stats(model,digits=22)
```

## Report One Way ANOVA with effect size power and post hocs
```{r report_one-way,echo=TRUE,message=FALSE,warning=FALSE,results='hide'}
result0<-report_oneway(df=mtcars,dv=2:4,iv=9:10,plot_diagnostics=FALSE,plot_means=FALSE)
result1<-report_oneway(df=mtcars,dv=2:4,iv=9:10,plot_diagnostics=TRUE,plot_means=TRUE)
```

```{r print_one-way, echo=FALSE}
result0
result1
```

## Report Factorial ANOVA with effect size power and post hocs
```{r report_factorial,echo=TRUE,warning=FALSE}
set.seed(12345)
df<-data.frame(id=rep(seq(1,80),each=81,1),
               IV1=rep(LETTERS[1:3],each=1,2160),
               IV2=rep(LETTERS[4:6],each=3,720),
               IV3=rep(LETTERS[7:9],each=9,240),
               IV4=rep(LETTERS[10:12],each=27,80),
               stringsAsFactors=FALSE)
cdf<-data.frame(matrix(.01,ncol=4,nrow=4))
correlation_martix<-as.matrix(cdf)
diag(correlation_martix)<-1
cdf<-generate_correlation_matrix(correlation_martix,nrows=nrow(df))+10
names(cdf)<-paste0("DV",1:4)
df<-data.frame(df,cdf)
df$DV2<-df$DV2+10
df$DV3<-df$DV3+20
df$DV4<-df$DV4+30
df[df$IV1%in%"A",]$DV1<-df[df$IV1%in%"A",]$DV1+1
df[df$IV1%in%"B",]$DV1<-df[df$IV1%in%"B",]$DV1+2
df[df$IV1%in%"C",]$DV1<-df[df$IV1%in%"C",]$DV1+3
cdf(df)
r1<-report_factorial_anova(df=df,
                           wid="id",
                           dv=c("DV1","DV2"),
                           within=c("IV1","IV2"),
                           within_full=c("IV1","IV2"),
                           between=NULL,
                           within_covariates=NULL,
                           between_covariates=NULL,
                           post_hoc=TRUE)
r2<-report_factorial_anova(df=df,
                           wid="id",
                           dv=c("DV1","DV2"),
                           within=NULL,
                           within_full=NULL,
                           between=c("IV1","IV2"),
                           within_covariates=NULL,
                           between_covariates=NULL,
                           post_hoc=TRUE)
r3<-report_factorial_anova(df=df,
                           wid="id",
                           dv=c("DV1","DV2"),
                           # within=c("IV3","IV4"),
                           # within_full=c("IV3","IV4"),
                           between=c("IV1","IV2"),
                           within_covariates=NULL,
                           between_covariates=NULL,
                           post_hoc=FALSE)
r4<-report_factorial_anova(df=df,
                           wid="id",
                           dv=c("DV1","DV2"),
                           within=c("IV1","IV2"),
                           within_full=c("IV1","IV2"),
                           between=NULL,
                           within_covariates=c("DV3","DV4"),
                           between_covariates=NULL,
                           post_hoc=TRUE)
r1
r2
r3
r4

```

## MANOVA
```{r report_manova,echo=TRUE}
## Set orthogonal contrasts.
op<-options(contrasts=c("contr.helmert","contr.poly"))
model_mixed<-manova(cbind(yield,foo)~N*P*K,within(npk,foo<-rnorm(24)))
model_between<-manova(cbind(rnorm(24),rnorm(24))~round(rnorm(24),0)*round(rnorm(24),0))
report_manova(model=model_mixed)
report_manova(model=model_between)
```

## Effect Size
```{r effect_size,echo=TRUE}
form<-formula(uptake~Treatment)
one_way_between<-aov(form,CO2)
factorial_between<-aov(uptake~Treatment*Type,CO2)
compute_aov_es(model=one_way_between,ss="I")
sjstats::anova_stats(one_way_between,digits=10)
compute_aov_es(model=one_way_between,ss="II")
sjstats::anova_stats(one_way_between,digits=10)
compute_aov_es(model=one_way_between,ss="III")
sjstats::anova_stats(one_way_between,digits=10)
compute_aov_es(model=factorial_between,ss="I")
sjstats::anova_stats(factorial_between,digits=10)
compute_aov_es(model=factorial_between,ss="II")
sjstats::anova_stats(factorial_between,digits=10)
compute_aov_es(model=factorial_between,ss="III")
sjstats::anova_stats(car::Anova(factorial_between,Type=3),digits=10)
```

## Post Hoc
```{r post_hoc,echo=TRUE}
compute_posthoc(mtcars[,6],mtcars[,10])
```
