---
title: "GLM MEANS"
author: "Dimitrios Zacharatos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GLM MEANS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
</style>

# Introduction
This shows the output of report_ttests and report_wtests functions from workingfunctions.

```{r setup, include = FALSE}
library(workingfunctions)
options(width = 1000)
```

# T TEST
```{r, report_ttests}
report_ttests(df=mtcars,dv=2,iv=9,alternative="two.sided")
report_ttests(df=mtcars,dv=2,iv=9:10,alternative="less")
report_ttests(df=mtcars,dv=2:3,iv=9,alternative="greater")
report_ttests(df=mtcars,dv=2:3,iv=9:10)
```

## EXPLANATION of COLUMNS
| Column   | Description                                                                 |
|----------|-----------------------------------------------------------------------------|
|DV |                 "dependent variable"  |
|IV |                 "independent variable"  |
|level1|              "level 1"  |
|level2|              "level 2"  |
|n1|                  "sample size for level 1"  |
|n2|                  "sample size for level 2"  |
|t|                   "t statistic"  |
|df|                  "degrees of freedom for t statistic"  |
|p|                   "p value"  |
|CI_l|                "confidence interval lower bound"  |
|CI_u|                "confidence interval upper bound"  |
|alternative|         "alternative hypothesis"  |
|method|              ""  |
|mean1|               "mean for level 1"  |
|mean2|               "mean for level 2"  |
|sd1|                 "standard deviation for level 1"  |
|sd2|                 "standard deviation for level 2"  |
|sd_pooled|           "pooled standard deviation\n\ncombined standard deviation of both levels"  |
|d|                   "measure of effect size\n\ncohen's d\n\nVery small\t0.01\t\tSawilowsky (2009)\nSmall\t\t0.20\t\tCohen (1988)\nMedium\t\t0.50\t\tCohen (1988)\nLarge\t\t0.80\t\tCohen (1988)\nVery large\t\t1.20\t\tSawilowsky (2009)\nHuge\t\t12.0\t\t1Sawilowsky (2009)"  |
|r|                   "measure of effect size\n\ncorrelation effect size"  |
|k_squared.bartlett.| "bartlett test for homogeneity of variances\n\n"  |
|df.bartlett.|        "bartlett test for homogeneity of variances\n\n"  |
|p.bartlett.|         "bartlett test for homogeneity of variances\n\nsignificant values indicate heteroscedasticity"  |
|bonferroni_p|        "bonferroni adjusted critical value for a=0.05"  |
|significant|         "if TRUE result is significant after bonferroni adjustment"  |

# WILCOXON TEST
```{r, report_wtests}
report_wtests(df=mtcars,dv=2,iv=9,alternative="two.sided",exact=FALSE)
report_wtests(df=mtcars,dv=2,iv=9:10,alternative="less",exact=FALSE)
report_wtests(df=mtcars,dv=2:3,iv=9,alternative="greater",exact=FALSE)
report_wtests(df=mtcars,dv=2:3,iv=9:10,exact=FALSE)
```


## EXPLANATION of COLUMNS

| Column   | Description                                                                 |
|----------|-----------------------------------------------------------------------------|
|DV                  |"dependent variable"  |
|IV                  |"independent variable"  |
|level1              |"level 1"  |
|level2              |"level 2"  |
|n1                  |"sample size for level 1"  |
|n2                  |"sample size for level 2"  |
|W                   |"wilcoxon statistic"  |
|df                  |"degrees of freedom for t statistic"  |
|p                   |"p value"  |
|CI_l                |"confidence interval lower bound"  |
|CI_u                |"confidence interval upper bound"  |
|alternative         |"alternative hypothesis"  |
|method              |""  |
|mean1               |"mean for level 1"  |
|mean2               |"mean for level 2"  |
|sd1                 |"standard deviation for level 1"  |
|sd2                 |"standard deviation for level 2"  |
|sd_pooled           |"pooled standard deviation\n\ncombined standard deviation of both levels"  |
|d                   |"measure of effect size\n\ncohen's d\n\nVery small\t0.01\t\tSawilowsky (2009)\nSmall\t\t0.20\t\tCohen (1988)\nMedium\t\t0.50\t\tCohen (1988)\nLarge\t\t0.80\t\tCohen (1988)\nVery large\t\t1.20\t\tSawilowsky (2009)\nHuge\t\t12.0\t\t1Sawilowsky (2009)"  |
|r                   |"measure of effect size\n\ncorrelation effect size"  |
|k_squared.bartlett. |"bartlett test for homogeneity of variances\n\n"  |
|df.bartlett.        |"bartlett test for homogeneity of variances\n\n"  |
|p.bartlett.         |"bartlett test for homogeneity of variances\n\nsignificant values indicate heteroscedasticity"  |
|bonferroni_p        |"bonferroni adjusted critical value for a=0.05"  |
|significant         |"if TRUE result is significant after bonferroni adjustment"  |
