---
title: "GLM_MEANS"
author: "Dimitrios Zacharatos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GLM_MEANS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style> 
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
/* Styling for input code */
pre { 
  background-color: #f4f4f4; /* Light gray background for input */
  border-left: 5px solid #007ACC; /* Blue border */
  padding: 10px;
}

/* Styling for output */
pre.r-output {
  background-color: #e8f5e9; /* Light green background for output */
  border-left: 5px solid #388E3C; /* Green border */
  padding: 10px;
}
</style>

This shows the output of report_ttests and report_wtests functions from workingfunctions.  
Installation instructions for workingfunctions can be found [here](https://github.com/sedzinfo/workingfunctions)

These functions perform t tests and wilcoxon tests for all possible combinations of levels of the independent variables.

```{r options,echo=FALSE,warning=FALSE,message=FALSE}
options(width=1000)
hook_output <- function(x, options) {
  paste0('<pre class="r-output">', knitr::knit_print(x), '</pre>')
}
knitr::knit_hooks$set(output = hook_output)
```

```{r load,warning=FALSE,message=FALSE}
library(workingfunctions)
```

# T TEST
```{r, report_ttests}
report_ttests(df=mtcars,dv=2,iv=9,alternative="two.sided")
report_ttests(df=mtcars,dv=2,iv=9:10,alternative="less")
report_ttests(df=mtcars,dv=2:3,iv=9,alternative="greater")
report_ttests(df=mtcars,dv=2:3,iv=9:10)
```

## Explanation of columns
| Column   | Description                                                                 |
|----------|-----------------------------------------------------------------------------|
|DV                  |dependent variable|
|IV                  |independent variable|
|level1              |level 1|
|level2              |level 2|
|n1                  |sample size for level 1|
|n2                  |sample size for level 2|
|t                   |t statistic|
|df                  |degrees of freedom for t statistic|
|p                   |p value|
|CI_l                |confidence interval lower bound|
|CI_u                |confidence interval upper bound|
|alternative         |alternative hypothesis|
|method              |name of method should be displayed here|
|mean1               |mean for level 1|
|mean2               |mean for level 2|
|sd1                 |standard deviation for level 1|
|sd2                 |standard deviation for level 2|
|sd_pooled           |pooled standard deviation combined standard deviation of both levels|
|d                   |measure of effect size cohen's d|
|d                   |Very small 0.01 Sawilowsky (2009)|
|d                   |Small 0.20 Cohen (1988)|
|d                   |Medium 0.50 Cohen (1988)|
|d                   |Large 0.80 Cohen (1988)|
|d                   |Very large 1.20 Sawilowsky (2009)|
|d                   |Huge 12.0 Sawilowsky (2009)|
|r                   |measure of effect size correlation effect size|
|k_squared bartlett  |bartlett test for homogeneity of variances|
|df bartlett         |bartlett test for homogeneity of variances|
|p bartlett          |bartlett test for homogeneity of variances significant values indicate heteroscedasticity|
|bonferroni_p        |bonferroni adjusted critical value for a=0.05|
|significant         |if TRUE result is significant after bonferroni adjustment|

# WILCOXON TEST
```{r, report_wtests}
report_wtests(df=mtcars,dv=2,iv=9,alternative="two.sided",exact=FALSE)
report_wtests(df=mtcars,dv=2,iv=9:10,alternative="less",exact=FALSE)
report_wtests(df=mtcars,dv=2:3,iv=9,alternative="greater",exact=FALSE)
report_wtests(df=mtcars,dv=2:3,iv=9:10,exact=FALSE)
```

## Explanation of columns
| Column   | Description                                                                 |
|----------|-----------------------------------------------------------------------------|
|DV                  |dependent variable|
|IV                  |independent variable|
|level1              |level 1|
|level2              |level 2|
|n1                  |sample size for level 1|
|n2                  |sample size for level 2|
|W                   |wilcoxon statistic|
|df                  |degrees of freedom for t statistic|
|p                   |p value|
|CI_l                |confidence interval lower bound|
|CI_u                |confidence interval upper bound|
|alternative         |alternative hypothesis|
|method              |name of method should be displayed here|
|mean1               |mean for level 1|
|mean2               |mean for level 2|
|sd1                 |standard deviation for level 1|
|sd2                 |standard deviation for level 2|
|sd_pooled           |pooled standard deviation combined standard deviation of both levels|
|d                   |measure of effect size cohen's d|
|d                   |Very small 0.01 Sawilowsky (2009)|
|d                   |Small 0.20 Cohen (1988)|
|d                   |Medium 0.50 Cohen (1988)|
|d                   |Large 0.80 Cohen (1988)|
|d                   |Very large 1.20 Sawilowsky (2009)|
|d                   |Huge 12.0 Sawilowsky (2009)|
|r                   |measure of effect size correlation effect size|
|k_squared bartlett  |bartlett test for homogeneity of variances|
|df bartlett         |bartlett test for homogeneity of variances|
|p bartlett          |bartlett test for homogeneity of variances significant values indicate heteroscedasticity|
|bonferroni_p        |bonferroni adjusted critical value for a=0.05|
|significant         |if TRUE result is significant after bonferroni adjustment|
