---
title: "HLR"
author: "Dimitrios Zacharatos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HLR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body { max-width: 95%; margin: 0 auto; padding: 0; } 
.container { width: 100%; max-width: none; } 
/* Example for wider tables */ 
table { width: 80%; table-layout: fixed; } 
th, td { padding: 8px; text-align: left; } 
/* Styling for input code */
pre { 
  background-color: #f4f4f4; /* Light gray background for input */
  border-left: 5px solid #007ACC; /* Blue border */
  padding: 10px;
  font-size: 20px !important;
}
/* Styling for output */
pre.r-output {
  background-color: #e8f5e9; /* Light green background for output */
  border-left: 5px solid #388E3C; /* Green border */
  padding: 10px;
  font-size: 15px !important;
}
</style>

```{r options,echo=FALSE,warning=FALSE,message=FALSE}
options(future.show.progress=FALSE)
options(width=1000)
knitr::opts_chunk$set(echo=TRUE)
hook_output<-function(x, options) {
  paste0('<pre class="r-output">',knitr::knit_print(x),'</pre>')
}
knitr::knit_hooks$set(output=hook_output)
```

```{r load,include=FALSE}
library(workingfunctions)
```

This shows the output of HLR function from workingfunctions.  
Installation instructions of workingfunctions can be found here https://github.com/sedzinfo/workingfunctions

## Report HLR
```{r clear_text}
result<-report_hlr(df=infert,corlist=8,factorlist=1,predictor="case",random_effect="case")
result
```
