---
title: "Confirmatory Factor Analysis"
subtitle: "Measurement Model & Covariance Structure"
output:
  ioslides_presentation:
    widescreen: true
    mathjax: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(lavaan)

y1 <- function(x) {-1 + 1*x}
y2 <- function(x) { 0 + 1*x}
y3 <- function(x) { 1 + 1*x}
y4 <- function(x) { 0 + -1*x}
y5 <- function(x) { 0 +  0*x}
y6 <- function(x) { 0 +  1*x}

xrange <- seq(-3, 3, 0.1)
df_data <- data.frame(
  x      = xrange,
  res_y1 = y1(xrange),
  res_y2 = y2(xrange),
  res_y3 = y3(xrange),
  res_y4 = y4(xrange),
  res_y5 = y5(xrange),
  res_y6 = y6(xrange)
)

df_long <- pivot_longer(df_data, cols = -x, names_to = "fn", values_to = "y")

meta <- data.frame(
  fn    = paste0("res_y", 1:6),
  label = c("τ=−1, λη=1", "τ=0, λη=1", "τ=1, λη=1",
            "τ=0, λη=−1", "τ=0, λη=0", "τ=0, λη=1"),
  row   = c(rep("λη = 1  (τ varies)", 3), rep("τ = 0  (λη varies)", 3))
)
# meta$label <- factor(meta$label, levels = unique(meta$label))
# meta$row   <- factor(meta$row,   levels = unique(meta$row))

df_long <- merge(df_long, meta, by = "fn")

plot_regression_lamnda<-ggplot(df_long[grep("(λη varies)",df_long$row),], aes(x = x, y = y)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.5) +
  geom_line(color = "steelblue", linewidth = 1.4) +
  facet_grid(row ~ label) +
  labs(x = expression(eta), y = "y") +
  coord_cartesian(xlim = c(-4, 4), ylim = c(-4, 4)) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0, hjust = 0)
  )

plot_regression_t<-ggplot(df_long[grep("(τ varies)",df_long$row),], aes(x = x, y = y)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.5) +
  geom_line(color = "steelblue", linewidth = 1.4) +
  facet_grid(row ~ label) +
  labs(x = expression(eta), y = "y") +
  coord_cartesian(xlim = c(-4, 4), ylim = c(-4, 4)) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0, hjust = 0)
  )


pop_model <- "
  eta =~ 1*y1 + 1*y2 + 1*y3
  y1 ~~ 1*y1
  y2 ~~ 1*y2
  y3 ~~ 1*y3
  eta ~~ 1*eta
"

# --- 2. simulate data ---
set.seed(42)
dat <- simulateData(pop_model)
# --- 3. fit CFA ---
cfa_model <- "
  eta =~ y1 + y2 + y3
"

fit <- cfa(cfa_model, data = dat)

# --- 4. results ---
# summary(fit, fit.measures = TRUE, standardized = TRUE)
# parameterEstimates(fit, standardized = TRUE)

plot_path <- lavaanPlot::lavaanPlot(
  model      = fit,
  node_options = list(shape = "box", fontname = "Helvetica"),
  edge_options = list(color = "grey"),
  coefs      = FALSE,   # show path coefficients
  stand      = TRUE,   # standardised
  stars      = "latent"
)

```

## Population & Sample Covariance

**Population covariance** between items $j$ and $k$:

$$\sigma_{jk} = \frac{1}{n}\sum_{i=1}^{n}(y_{ij} - \bar{y}_j)(y_{ik} - \bar{y}_k)$$

**Sample covariance** (estimating $\sigma_{jk}$ from $n$ observations):

$$s_{jk} = \frac{1}{n-1}\sum_{i=1}^{n}(y_{ij} - \bar{y}_j)(y_{ik} - \bar{y}_k)$$

## Estimating Covariance: Worked Example 1

$n = 4$ observations, $\bar{x} = 5$, $\bar{y} = 5$:

| $i$ | $x_i$ | $y_i$ | $x_i - \bar{x}$ | $y_i - \bar{y}$ | $(x_i-\bar{x})(y_i-\bar{y})$ |
|:---:|:------:|:------:|:---------------:|:---------------:|:-----------------------------:|
| 1 | 2 | 4 | $-3$ | $-1$ | $3$ |
| 2 | 4 | 5 | $-1$ | $0$ | $0$ |
| 3 | 6 | 5 | $1$ | $0$ | $0$ |
| 4 | 8 | 6 | $3$ | $1$ | $3$ |
| | | | | **Sum** | $\mathbf{6}$ |

## Estimating Covariance: Worked Example 2

$$s_{xy} = \frac{1}{n-1}\sum_{i=1}^{4}(x_i - \bar{x})(y_i - \bar{y}) = \frac{6}{4-1} = \frac{6}{3} = 2$$

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=6}

cov(c(2,4,6,8), c(4,5,5,6))

```

> On average, when $x$ is 1 unit above its mean, $y$ tends to be **2 units** above its mean.

## Correlation
**Correlation** — covariance scaled by the product of standard deviations:

$$\rho_{jk} = \frac{E\!\left[(y_j - \mu_j)(y_k - \mu_k)\right]}{\sqrt{E\!\left[(y_j-\mu_j)^2\right]\cdot E\!\left[(y_k-\mu_k)^2\right]}} = \frac{\sigma_{jk}}{\sigma_j \sigma_k}$$

$$r_{jk} = \frac{\displaystyle\sum_{i=1}^{n}(y_{ij}-\bar{y}_j)(y_{ik}-\bar{y}_k)}{\displaystyle\sqrt{\sum_{i=1}^{n}(y_{ij}-\bar{y}_j)^2 \;\cdot\; \sum_{i=1}^{n}(y_{ik}-\bar{y}_k)^2}}$$

## Correlation Covariance comparison
| | Covariance | Correlation |
|---|---|---|
| Scale | Unbounded | $-1 \leq r \leq 1$ |
| Units | Product of item units | Unitless |
| Diagonal | $\sigma^2_j$ (variance) | $1$ |

> **Goal of CFA:** find $\theta$ such that $\boldsymbol{\Sigma}(\theta) \approx \mathbf{S}$

## Measurement Model — Scalar Form

For a 3 item one factor case,
each observed item $y$ is modeled as:

$$
\begin{matrix}
y_1 = \tau_1 + \lambda_{1}\eta_{1} + \epsilon_{1} \\
y_2 = \tau_2 + \lambda_{2}\eta_{1} + \epsilon_{2} \\
y_3 = \tau_3 + \lambda_{3}\eta_{1} + \epsilon_{3}
\end{matrix}
$$

| Symbol | Name | Interpretation |
|--------|------|----------------|
| $\tau$ | Intercepts | Item means |
| $\lambda$ | Loadings | Correlation of item with factor |
| $\eta$ | Latent factor | The underlying construct (e.g. SPSS Anxiety) |
| $\epsilon$ | Residuals | What the factor does **not** explain |


## Effect of $\tau$
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=6}
plot_regression_t
```

## Effect of $\lambda$
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=6}
plot_regression_lamnda
```

## Measurement Model — Matrix Form

For a 3 item one factor case:
$$
\begin{matrix}
y_1 = \tau_1 + \lambda_{1}\eta_{1} + \epsilon_{1} \\
y_2 = \tau_2 + \lambda_{2}\eta_{1} + \epsilon_{2} \\
y_3 = \tau_3 + \lambda_{3}\eta_{1} + \epsilon_{3}
\end{matrix}
$$

For a 3 item one factor case:
$$
\begin{pmatrix} y_{1} \\ y_{2} \\ y_{3} \end{pmatrix}
=
\begin{pmatrix} \tau_1 \\ \tau_2 \\ \tau_3 \end{pmatrix}
+
\begin{pmatrix} \lambda_{1} \\ \lambda_{2} \\ \lambda_{3} \end{pmatrix}
\begin{pmatrix} \eta_{1} \end{pmatrix}
+
\begin{pmatrix} \epsilon_{1} \\ \epsilon_{2} \\ \epsilon_{3} \end{pmatrix}
$$

That is:
$$\mathbf{y} = \boldsymbol{\tau} + \boldsymbol{\Lambda}\boldsymbol{\eta} + \boldsymbol{\epsilon}$$

## Path diagram

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=5}

plot_path

```

## Covariance Structure

The **model-implied covariance matrix** $\Sigma(\theta)$ is:

$$\Sigma(\theta) = \mathbf{\Lambda \Psi \Lambda'} + \Theta_{\epsilon}$$

| Symbol | Name | Description |
|--------|------|-------------|
| $\Lambda$ | Factor loading matrix | Same $\lambda$'s from the measurement model |
| $\Psi$ | Latent covariance matrix | Variance of $\eta$; scalar for one factor |
| $\Theta_{\epsilon}$ | Residual covariance matrix | Variance-covariance of residuals |

## Covariance Structure — Expanded

$$
\Sigma(\theta) =
\begin{pmatrix} \lambda_{1} \\ \lambda_{2} \\ \lambda_{3} \end{pmatrix}
\begin{pmatrix} \psi_{11} \end{pmatrix}
\begin{pmatrix} \lambda_{1} & \lambda_{2} & \lambda_{3} \end{pmatrix}
+
\begin{pmatrix}
\theta_{11} & \theta_{12} & \theta_{13} \\
\theta_{21} & \theta_{22} & \theta_{23} \\
\theta_{31} & \theta_{32} & \theta_{33}
\end{pmatrix}
$$

- $\mathbf{\Lambda \Psi \Lambda’}$ captures the **shared variance** due to the factor
- $\Theta_{\epsilon}$ captures the **unique variance** of each item

$$\Sigma(\theta) = \text{Cov}(\mathbf{y}) = \boldsymbol{\Lambda} \Psi \mathbf{\Lambda}’ + \Theta_{\epsilon}$$


## Model Degrees of Freedom — The Logic

$$df = \underbrace{\frac{p(p+1)}{2}}_{\text{known}} \;-\; \underbrace{q}_{\text{free parameters}}$$

**Known quantities** = unique elements of the sample covariance matrix $\mathbf{S}$

For $p$ observed variables: $\frac{p(p+1)}{2}$ values (lower triangle + diagonal)

$$\mathbf{S} = \begin{pmatrix} s_{11} & & \\ s_{21} & s_{22} & \\ s_{31} & s_{32} & s_{33} \end{pmatrix} \quad \Rightarrow \quad \frac{3 \cdot 4}{2} = 6 \text{ unique values}$$

**Free parameters** $q$ = everything CFA must estimate from those values

## Model Degrees of Freedom — The Logic cont
  
$\quad df > 0$ → model is **over-identified** (testable)  
$\quad df = 0$ → model is **just-identified** (no test possible)  
$\quad df < 0$ → model is **under-identified** (cannot be estimated)  

## Model Degrees of Freedom — Counting Parameters

For a **one-factor CFA** with $p$ items (marker-variable identification: $\lambda_1 = 1$):

| Parameter | Symbol | Free count | Reason |
|---|---|:---:|---|
| Factor loadings | $\lambda_2, \ldots, \lambda_p$ | $p - 1$ | $\lambda_1$ fixed to 1 |
| Factor variance | $\psi_{11}$ | $1$ | Scale of $\eta$ |
| Residual variances | $\theta_{\epsilon_1}, \ldots, \theta_{\epsilon_p}$ | $p$ | One per item |
| **Total** $q$ | | $\mathbf{2p}$ | |

$$df = \frac{p(p+1)}{2} - 2p = \frac{p(p-3)}{2}$$

## Model Degrees of Freedom — Counting Parameters cont

| $p$ | Knowns | Free $q$ | $df$ | Status |
|:---:|:---:|:---:|:---:|:---:|
| 3 | 6 | 6 | **0** | just-identified |
| 4 | 10 | 8 | **2** | over-identified |
| 5 | 15 | 10 | **5** | over-identified |
| 6 | 21 | 12 | **9** | over-identified |

> Each additional item beyond 3 adds $p - 2$ new $df$ — more items = richer test of fit.

## Types of Model Fit Indices

**Absolute fit** — how well does $\hat{\boldsymbol{\Sigma}}(\theta)$ reproduce the observed $\mathbf{S}$?

> Compares the model directly to the data, with no reference to any other model. A value of zero means perfect reproduction.

**Incremental fit** (Comparative) — how much better is the target model than the worst possible model?

> Scales fit relative to a **baseline (null) model** in which all observed variables are uncorrelated ($\boldsymbol{\Sigma}_b = \text{diag}(\mathbf{S})$). A value of 1 means the target model explains all the covariance structure the null model cannot.

**Parsimony fit** — does the model fit well *given its complexity*?

> Penalises models for having many free parameters ($q$). A model that fits perfectly but uses every available degree of freedom is not parsimonious. Parsimony indices reward simpler explanations.

| Family | Logic | Representative indices |
|---|---|---|
| Absolute | $\hat{\boldsymbol{\Sigma}} \approx \mathbf{S}$? | $\chi^2$, SRMR, GFI |
| Incremental | Better than null? | CFI, TLI, NFI |
| Parsimony | Fit per $df$? | RMSEA, AIC, BIC |

## Model Fit Indices 1

| Type | Index | Formula / Benchmark | Acceptable |
|------|-------|---------------------|------------|
| **Absolute** | $\chi^2$ | $\sum w_{jk}(\hat{\sigma}_{jk} - s_{jk})^2$ | $p > .05$ (sensitive to $n$) |
| **Absolute** | SRMR | $\sqrt{\frac{2\sum_j\sum_{k\leq j}(s_{jk}-\hat{\sigma}_{jk})^2}{p(p+1)}}$ | $< .08$ |
| **Absolute** | GFI | $1 - \frac{\hat{F}}{F_b}$ | $> .90$ |
| **Incremental** | CFI | $1 - \frac{\hat{F} - df_t}{F_b - df_b}$ | $> .95$ |
| **Incremental** | TLI / NNFI | $\frac{\chi^2_b/df_b \;-\; \chi^2_t/df_t}{\chi^2_b/df_b - 1}$ | $> .95$ |

> Subscript $b$ = baseline (null) model; $t$ = target model; $p$ = number of observed variables

## Model Fit Indices 2
| Type | Index | Formula / Benchmark | Acceptable |
|------|-------|---------------------|------------|
| **Incremental** | NFI | $\frac{\chi^2_b - \chi^2_t}{\chi^2_b}$ | $> .90$ |
| **Parsimony** | RMSEA | $\sqrt{\max\!\left(0,\frac{\chi^2 - df}{df(n-1)}\right)}$ | $< .06$ |
| **Parsimony** | AIC | $\chi^2 - 2\,df$ | Lower is better |
| **Parsimony** | BIC | $\chi^2 - df\ln(n)$ | Lower is better |

> Subscript $b$ = baseline (null) model; $t$ = target model; $p$ = number of observed variables


## Fit Index: $\chi^2$ — Chi-Square Test | *Absolute*

$$\chi^2 = (n-1)\, F(\mathbf{S},\, \hat{\boldsymbol{\Sigma}}) \qquad df = \tfrac{p(p+1)}{2} - q$$

- Tests whether $\boldsymbol{\Sigma}(\theta) = \mathbf{S}$ exactly in the population
- Assess overall fit and the discrepancy between the sample and fitted covariance matrices. 
- **$H_0$**: the model fits perfectly — so a **non-significant** $p$ is desired
- **Problem**: with large $n$, trivial misfit becomes significant; with small $n$, poor models pass
- Best used as a relative comparison between nested models ($\Delta\chi^2$ test)

## Fit Index: $\chi^2$ — Chi-Square Test | *Absolute* cont
| | |
|---|---|
| Acceptable | $p > .05$ |
| Sensitive to | Sample size $n$, non-normality |

## Fit Index: SRMR — Standardized Root Mean Square Residual | *Absolute*

$$\text{SRMR} = \sqrt{\frac{2\sum_j\sum_{k \leq j}(s_{jk} - \hat{\sigma}_{jk})^2}{p(p+1)}}$$

- Average discrepancy between observed and model-implied **correlations**
- Standardizing makes it comparable across studies with different item scales
- Value of **0** = perfect fit; increases as residuals grow
- Sensitive to misspecified **factor loadings** (less so to structural paths)

## Fit Index: SRMR — Standardized Root Mean Square Residual | *Absolute* cont

| | |
|---|---|
| Acceptable | $< .08$ |
| Good | $< .05$ |

## Fit Index: GFI — Goodness of Fit Index | *Absolute*

$$\text{GFI} = 1 - \frac{\hat{F}(\mathbf{S},\,\hat{\boldsymbol{\Sigma}})}{F(\mathbf{S},\,\mathbf{I})}$$

- Proportion of variance in $\mathbf{S}$ explained by $\hat{\boldsymbol{\Sigma}}$; analogous to $R^2$
- Ranges $[0, 1]$; higher = better fit
- **Limitation**: inflated by large $n$ and number of parameters; not recommended as a primary index in modern practice
- AGFI (Adjusted GFI) penalises for degrees of freedom: $\text{AGFI} = 1 - \frac{p(p+1)}{2\,df}(1-\text{GFI})$

## Fit Index: GFI — Goodness of Fit Index | *Absolute* cont
| | |
|---|---|
| Acceptable | $> .90$ |
| Note | Superseded by CFI/RMSEA in most guidelines |

## Fit Index: CFI — Comparative Fit Index | *Incremental*

$$\text{CFI} = 1 - \frac{\max(\chi^2_t - df_t,\; 0)}{\max(\chi^2_b - df_b,\; 0)}$$

- Compares the target model to the **null model** (all covariances $= 0$)
- Corrects for the positive bias of NFI with small samples
- Ranges $[0, 1]$; robust to sample size
- Most widely reported incremental index in CFA/SEM

## Fit Index: CFI — Comparative Fit Index | *Incremental* cont
| | |
|---|---|
| Acceptable | $> .95$ (Hu & Bentler, 1999) |
| Adequate | $> .90$ |
| Subscripts | $t$ = target model, $b$ = baseline (null) model |

## Fit Index: TLI / NNFI — Tucker–Lewis Index | *Incremental*

$$\text{TLI} = \frac{\chi^2_b / df_b \;-\; \chi^2_t / df_t}{\chi^2_b / df_b \;-\; 1}$$

- Also called **Non-Normed Fit Index (NNFI)**
- Penalises model complexity: divides $\chi^2$ by $df$ before comparing
- Can exceed 1 or fall below 0 for very good or very poor models
- Preferred over NFI when comparing models of different complexity

## Fit Index: TLI / NNFI — Tucker–Lewis Index | *Incremental* cont
| | |
|---|---|
| Acceptable | $> .95$ |
| Adequate | $> .90$ |
| Note | Values $> 1$ are truncated to 1 in most software |

## Fit Index: NFI — Normed Fit Index | *Incremental*

$$\text{NFI} = \frac{\chi^2_b - \chi^2_t}{\chi^2_b}$$

- Proportion of the null model's $\chi^2$ eliminated by the target model
- First incremental index proposed (Bentler & Bonett, 1980)
- Ranges strictly $[0, 1]$ — hence "normed"
- **Limitation**: underestimates fit with small $n$; does not penalise for complexity

## Fit Index: NFI — Normed Fit Index | *Incremental* cont
| | |
|---|---|
| Acceptable | $> .90$ |
| Limitation | Biased downward when $n < 200$ |
| Superseded by | CFI, TLI for most applications |

## Fit Index: RMSEA — Root Mean Square Error of Approximation | *Parsimony*

$$\text{RMSEA} = \sqrt{\max\!\left(0,\; \frac{\chi^2 - df}{df\,(n-1)}\right)}$$

- Measures **misfit per degree of freedom** — rewards parsimony
- Acknowledges that no real-world model fits perfectly; tests "close fit"
- 90% confidence interval routinely reported; test $H_0\!: \text{RMSEA} \leq .05$
- Tends to favour models with many parameters (large $df$)

## Fit Index: RMSEA — Root Mean Square Error of Approximation | *Parsimony* cont
| | |
|---|---|
| Good | $< .05$ |
| Acceptable | $\leq .08$ |
| Poor | $> .10$ |

## Fit Index: AIC — Akaike Information Criterion | *Parsimony*

$$\text{AIC} = \chi^2 - 2\,df$$

- Penalises fit function by the number of **free parameters** ($q$); $df$ decreases as $q$ increases
- Derived from information theory (Kullback–Leibler divergence)
- Has no absolute cutoff — used for **model comparison only**
- Smaller AIC = better balance of fit and parsimony

## Fit Index: AIC — Akaike Information Criterion | *Parsimony* cont
| | |
|---|---|
| Use | Compare non-nested models |
| Rule | $\Delta\text{AIC} > 10$ = strong preference for lower model |
| Note | Does not correct for sample size — prefer BIC with large $n$ |

## Fit Index: BIC — Bayesian Information Criterion | *Parsimony*

$$\text{BIC} = \chi^2 - df\,\ln(n)$$

- Penalises more heavily than AIC when $n > e^2 \approx 7.4$ (always in practice)
- Approximates the log Bayes factor between models
- Favours **simpler** models more strongly than AIC as $n$ grows
- Preferred for **model selection** when sample size is large


## Fit Index: BIC — Bayesian Information Criterion | *Parsimony* cont
| | |
|---|---|
| Use | Compare non-nested models |
| Rule | $\Delta\text{BIC} > 10$ = strong evidence for lower model |
| vs AIC | BIC penalises complexity more; AIC penalises misfit more |

